Eval expressions

### 2021-01-19 analysis/expressions.Rmd
Canonical expression

arithm_folded <- arithm %>% filter(map_lgl(expr_resolved, function(exp) {
  ast <- NA
  try(ast <- parse(text = exp)[[1]], silent = TRUE)
  # some expr_resolved have been truncated so we mark them as FALSE (even though they could be true)
  if (!is.na(ast)) {
    return(is_arithm_expr(ast))
  }
  else {
    return(FALSE)
  }
}))
```

@@

@@

>>>>

arithm_folded <- arithm %>% filter(map_lgl(expr_resolved, function(exp) {
  ast <- NA
  try(ast <- parse(text = exp)[[1]], silent = TRUE)
  # some expr_resolved have been truncated so we mark them as FALSE (even though they could be true)
  if (length(ast) > 1 || !is.na(ast)) {
    return(is_arithm_expr(ast))
  }
  else {
    return(FALSE)
  }
}))
```

@@

```{r}
with_formula <- E %>%
  filter(str_detect(expr_resolved, fixed("~")))
```

```{r}
n_with_formula <- with_formula %>%
  pull(src_ref) %>%
  n_distinct()
```

The total number of call sites with formulas is `r n_with_formula`.

```{r}
with_formula %>%
  group_by(expr_resolved) %>%
  summarise(n_call_sites = n_distinct(src_ref), n_packages = n_distinct(ev_package)) %>%
  datatable()
```

`~` is mostly used within a call to `model.frame`.

@@

# Better normalization of resolved expressions

We fold all constant arithmetic expressions and replace all variable names by a placeholder.

```{r}
canonic_expr <- function(exp) {
  if (is.call(exp)) {
    function_name <- exp[[1]]
    function_args <- exp[-1]
    res <- map(function_args, canonic_expr)
    # Function_name can be larger than one is a function with a namespace stats::model or another higher-order function
    # For instance, (function(x, y) x + y)(3, 4)
    if (length(function_name) == 1 && as.character(function_name) %in% arith_func && every(res, function(v) {
      v == "NUM"
    })) {
      return("NUM")
    }
    else if (length(function_name) == 1 && function_name == "::") {
      str_c(as.character(function_args[[1]]), "::", as.character(function_args[[2]]))
    }
    else {
      function_name <- if (length(function_name) == 1) {
        as.character(function_name)
      }
      else {
        canonic_expr(function_name)
      }
      return(str_c(str_c(function_name, collapse = ", "), "(", str_c(res, collapse = ", "), ")"))
    }
  }
  else if (is.symbol(exp)) {
    return("VAR")
  }
  else if (is.expression(exp)) {
    return(str_c(map(exp, canonic_expr), collapse = ", "))
  }
  else if (typeof(exp) %in% c("integer", "double", "complex")) {
    return("NUM")
  }
  else {
    return(deparse1(exp))
  }
}

canonic_expr_str <- function(exp) {
  ast <- NA
  try(ast <- parse(text = exp)[[1]], silent = TRUE)
  # some expr_resolved have been truncated so we mark them as FALSE (even though they could be true)
  if (is.symbol(ast) || is.language(ast) || length(ast) > 1 || !is.na(ast)) {
    return(canonic_expr(ast))
  }
  else {
    return("CANONIZATION ERROR")
  }
}
```

For performance reason, we only apply this function to the unique resolved expressions.

```{r}
res_exprs <- E %>%
  select(expr_resolved) %>%
  unique()

res_exprs <- res_exprs %>% mutate(expr_canonic = map_chr(expr_resolved, canonic_expr_str))
```
